---
title: "Methods of dealing big data"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### [data.table](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html)

  * concise and consistent syntax irrespective of the set of operations you would like to perform to achieve your end goal.
  
  * performing analysis fluidly without the cognitive burden of having to map each operation to a particular function from a set of functions available before to perform the analysis.
  
  * automatically optimising operations internally, and very effectively, by knowing precisely the data required for each operation and therefore very fast and memory efficient.
  
```{r}
library(data.table)
```

```{r}
flights <- fread("../dataSet/flight14.csv")
head(flights, 4)
dim(flights)
```
an enhanced version of data.frames. 
```{r}
DT <- data.table(ID = c("b","b","b","a","a","c"), a = 1:6, b = 7:12, c = 13:18)
DT
```

### subset rows and cols
```{r}
# rows
ans <- flights[origin == "JFK" & month == 3]
head(ans)
ans <- flights[1:2]
ans

# cols vector 
ans <- flights[, arr_delay]
head(ans)

# cols data.table 
ans <- flights[, list(arr_delay)]
head(ans)

# alternatively 
ans <- flights[, .(arr_delay, dep_delay)] # ans <- flights[, .(arr_delay, dep_delay)]
head(ans) 
```


