---
title: "Loan Prediction – Using PCA and Naive Bayes Classification "
output:
  html_document: default
---
Naive Bayes classification
==========================

Description 
-----------

Nowadays, there are numerous risks related to bank loans both for the banks and the borrowers, who get the loans. The risk analysis about bank loans needs understanding about the risk and the risk level. Banks need to analyze their customers for loan eligibility so that they can specifically target those customers. 

Banks wanted to automate the loan eligibility process (real time) based on customer details such as Gender, Marital Status, Age, Occupation, Income, debts and others, provided in their online application form. As the number of transactions in banking sector is rapidly growing and huge data volumes are available, the customers’ behavior can be easily analyzed and the risks around loan can be reduced. So, it is very important to predict the loan type and loan amount based on the banks’ data.

Data Description

Customer loan dataset has samples of about 100+ unique customer details, where each customer is represented in a unique row. 

Visit our blog post for further details about use cases, http://www.treselle.com/blog/


### **[Reference](https://github.com/treselle-systems/loan_prediction_using_PCA_and_naive_bayes_classification_with_R)**


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caTools)
library(e1071)
library(caret)
library(ElemStatLearn)
```

### load data 
```{r}
customer_loan_details <- read.csv("../dataSet/cutomer_loan_details.csv", sep = ",")
```

### function 
```{r}
# Print the structure of the dataframe
head(customer_loan_details, 4)

# Check for the NA values
any(is.na(customer_loan_details))

# Calculating DTI
customer_loan_details$dti <- (customer_loan_details$debts/customer_loan_details$income)*100

# Create loan_decision_status variable which is our target variable to use for loan prediction
customer_loan_details$loan_decision_status <- ifelse(customer_loan_details$loan_decision_type == 'Denied', 0, 1)

# Encoding the target variable as factor
customer_loan_details$loan_decision_status <- factor(customer_loan_details$loan_decision_status, levels = c(0, 1))

#Selecting the required fields for prediction
customer_loan_refined <- customer_loan_details[,c(3,4,6:8,11,13:14)]
head(customer_loan_refined, 4)

# Encoding the categorical variable as factors
customer_loan_refined$gender <- as.numeric(factor(customer_loan_refined$gender,
                                                  levels = c('Male','Female'),
                                                  labels = c(1,2)))

customer_loan_refined$marital_status <- as.numeric(factor(customer_loan_refined$marital_status,
                                                          levels = c('Divorced','Married','Single'),
                                                          labels = c(1,2,3)))

customer_loan_refined$occupation <- as.numeric(factor(customer_loan_refined$occupation,
                                                      levels = c('Accout','Business','IT','Manager','NYPD'),
                                                      labels = c(1,2,3,4,5)))

customer_loan_refined$loan_type <- as.numeric(factor(customer_loan_refined$loan_type,
                                                     levels = c('Auto','Credit','Home','Personal'),
                                                     labels = c(1,2,3,4)))

head(customer_loan_refined, 4)
```

### Splitting the customer_loan_refined dataset into training and test sets
```{r}
set.seed(123)
split = sample.split(customer_loan_refined$loan_decision_status, SplitRatio = 0.70)
training_set = subset(customer_loan_refined, split == TRUE)
test_set = subset(customer_loan_refined, split == FALSE)

#Applying Feature Scaling
training_set[-8] = scale(training_set[-8])
test_set[-8] = scale(test_set[-8])

head(training_set, 4)
```


### Applying Dimensionality reduction using PCA to training and test sets
```{r}
pca <- preProcess(x = training_set[-8], method = 'pca', pcaComp = 2)
training_set_pca <- predict(pca, training_set)
training_set_pca <- training_set_pca[c(2, 3, 1)]
test_set_pca <- predict(pca, test_set)
test_set_pca <- test_set_pca[c(2, 3, 1)]
head(test_set_pca, 4)
```

### Appling Naive Bayes classification model to predict the loan
```{r, include=FALSE}
classifier <- naiveBayes(x = training_set_pca[-3], y = training_set_pca$loan_decision_status)

# Predicting the Test set results
y_pred <- predict(classifier, newdata = test_set_pca[-3])

# confusionMatrix to calculate accuracy
confusionMatrix(table(test_set_pca[, 3], y_pred))
```


### Visualising the Test set results
```{r}
set <- test_set_pca

# Built the grid using X1, X2 by taking min-1 and max+1 values from test set
X1 <- seq(min(set[, 1]) - 1, max(set[, 1]) + 1, by = 0.01)
X2 <- seq(min(set[, 2]) - 1, max(set[, 2]) + 1, by = 0.01)
grid_set <- expand.grid(X1, X2)
colnames(grid_set) <- c('PC1', 'PC2')

# Predict the test set observations 
y_grid <- predict(classifier, newdata = grid_set)
```


### Plot the graph using actual observations from test set and predicted results
```{r}
plot(set[, -3], main = 'Naive Bayes (Test set)',
     xlab = 'PC1', ylab = 'PC2',
     xlim = range(X1), ylim = range(X2))
contour(X1, X2, matrix(as.numeric(y_grid), length(X1), length(X2)), add = TRUE)
points(grid_set, pch = '.', col = ifelse(y_grid == 1, 'springgreen3', 'tomato'))
points(set, pch = 21, bg = ifelse(set[, 3] == 1, 'green4', 'red3'))
```